---
title: "Day 2 afternoon computational"
format: html
editor: visual
execute: 
  warning: false
  message: false
---

```{r}
library(tidyverse)
```

## More indexing

```{r}
#----For loops laying the foundation 

cars <- c("blue", "orange", "white")
cars[1]

i <- seq(1:2)
cars [i]


```

You can assign new values based on position

```{r}
cars[3] <- "yellow"
cars
```

### More Matrices

```{r}
#-----wrap a vector into a matrix by rows and column 
#-----byrow argument being false means that it will wrap by column 1, then column 2 

fish_size <- matrix(c(34, 86, 99, 55), ncol = 2, nrow = 2, byrow = FALSE)
fish_size

class(fish_size)

#----- Example indexing 
fish_size[1, 2]
fish_size[2,2]

#-----check the type ... expecting double 
typeof(fish_size)

```

what happens if we create a matrix with different types of data?

```{r}

fish_type <- matrix(c("trout", 86, 99, 55L), ncol = 2, nrow = 2, byrow = FALSE)
fish_type

#-----numerics have now been coerced to characters 
class(fish_type)
typeof(fish_type)

#-----[2,0] means take 2nd row but keep all columns 
fish_size[2,]
#----- returns all columns but filters by second row

#-----now let's return all rows but filter by column 
fish_size[,1]

```

### Lists

reminder: A list combines vectors that can be of different length and type

```{r}

urchins <- list("blue", c(1, 2, 3), c("a cat", "a dog"), 5L)

urchins

#----- how to access the different components: use the tricky []
#----- the first ex will just regurgitate the list that is within the list, however the latter will give you whats in the list, which are characters ... I think that i like the $ method better

ex1 <- urchins[3]

class(ex1)

ex2 <- urchins[[3]]

class(ex2)

```

### Make a named list

```{r}

tacos <- list(topping = c("onion", "cilantro", "guacamole"), 
              filling = c("beans", "meat", "veggies"), 
              price = c(6.75, 8.25, 9.50))

#-----notice this convention: use = within function and assignment operator to store as a final object outside of a function 

tacos

#----using the dollar sign to index by name!

tacos$filling[3]
```

### T-test on significance between two groups

do i have enough evidence to make me think that these are taken from two populations with different means... null hypoth is that they come from the same population (same mean)

```{r}

a <- c(1, 2, 3)

b <- c(4,5,6)

my_ttest <- t.test(a, b)
my_ttest

#----- this is great: index by just one of the output values 

my_ttest$p.value


```

## use in-line code instead

#### The second value of the vector 'filling' within the list 'tacos' is `r tacos$filling[2]`

the above did not work ... look it up later

## Put that list 'tacos' into a dataframe

```{r}
topping <- c("onion", "cilantro", "guacamole")
filling <- c ("beans", "meat", "veggies")
price <- c(6.75, 8.25, 9.50)

df_1 <- data.frame(topping, filling, price)
df_1
```

Now if you check the class of this, it returns data frame... where the last ones were list class

Can also make a data frame like this:

```{r}

df_ex2 <- data.frame(topping = c("yeah", "oh", "sure"), filling = c("water", "mice", "pig"), price = c(1, 2, 3))
df_ex2

#-----or you could even do this ~~~called a tribble
#------allows you to write out the info in the same organization that you would expect to see it in the dataframe
#when you call it back it also returns some other information

df_ex_3 <- tribble(
  ~topping, ~filling, ~price, 
  "avo", "steak", 4, 
  "ranch", "chicken", 5
)
df_ex_3

#----- and you can still index it just the same 

df_ex_3$topping[2]
```
